Tail Indices and Scale Parameters in Financial Time Series

### 2017-03-21 papers/FX/draft.tex
Some revisions.
There seems to be a jumping behaviour in the optimal phi when X is t-distributed...

\subsection{Pareto-distributed equity return}
In this section, we look into the situation when $X$ has a shifted
Pareto distribution. Assume
\begin{equation}
  \label{eq:X_distr}
  F(x) = \left\{
  \begin{array}{ll}
    p \left(
    {K \over K - x}
    \right)^\alpha & x \leq 0 \\
    1 - (1 - p) \left(
    {K' \over K' + x}
    \right)^\beta & x > 0
  \end{array}
  \right.
\end{equation}
Assume $\alpha, \beta > 1$, $0 < p < 1$ and $K > 0$.
We have
\begin{eqnarray}
  && \mathcal U F (\alpha, \beta, K, K', \phi) \nonumber \\
  &=&
  \alpha K^\alpha  p
  \int_{-\infty}^0
  u\left[ (1 - \phi) e^r + \phi e^x \right]
  {1 + b \over (K - x)^{\alpha + 1}} dx
  \nonumber \\
  &&
  \beta K'^\beta (1 - p)
  \int_{0}^\infty
  u\left[ (1 - \phi) e^r + \phi e^x \right]
  {1 + b \1{x < q} \over (K' + x)^{\beta + 1}} dx 
  + \text{const.}
  % &&
  % - b u(\delta v) \left[
  %   1 - (1 - p) {
  %     K'^\beta
  %     \over
  %     (K' + q)^\beta
  %   }
  % \right]
  \label{eq:xxie1.0}
\end{eqnarray}
where the last term is a constant independent of $\phi$, $\alpha$ and $K$.
\begin{theorem}
  \label{thrm:I}
  Suppose functional $G(\cdot)$ is given as \eqref{eq:G_functional},
  function $u(\cdot)$ is increasing and differentiable, then
  \[
  \pd{G(F)}{\alpha} > 0,
  \quad
  \pd{G(F)}{K} < 0
  \]
\end{theorem}
Since $u(v)$ given in \eqref{eq:xxie1.0} as a
functional of the distribution function of $X$ is increasing with the
lower tail index $\alpha$ and decreasing with $K$, there is a curve of
equal preference on the plane of $(\alpha, K)$. Moving along this
curve in the direction of increasing $\alpha$, one expects the values
of $K$ to increase too, i.e. the estimated values of $\alpha$ and $K$
should appear positively correlated. This is indeed the case in some
real stock return data, e.g. the ``Energy'' and ``Information Technology''
sectors of S\&P 500 companies. This is illustrated in figures
\ref{fig:Energy_alpha_K} and \ref{fig:Information_Technology_alpha_K}.
The confidence bands of the Hill estimates of $\alpha$ of these
sectors are shown in figures \ref{fig:Energy_lower},
\ref{fig:Consumer_Staples_lower} and
\ref{fig:Information_Technology_lower}.

\begin{proof}
  Let
  \begin{equation}
    \label{eq:xxie2}
    \hat \phi := \argmax_{0 < \phi \leq 1} \mathcal U(F, \phi)
  \end{equation}
  We have
  \[
  G(F) = \mathcal U(F, \hat\phi)
  \]
  It follows
  \begin{eqnarray}
    \pd{G(F)}{\alpha}
    &=&
    \pd{\mathcal U(F, \phi)}{\alpha}(\alpha, K, \hat\phi)
    + \pd{\mathcal U(F, \phi)}{\phi}(\alpha, K, \hat\phi)
    \pd{\hat\phi}{\alpha}(\alpha, K)
    \label{eq:xxie3}
  \end{eqnarray}
  The definition \eqref{eq:xxie2} implies $\forall \alpha > 1$
  \begin{equation}
    \label{eq:xxie4}
    \pd{\mathcal U(F, \phi)}{\phi}(\alpha, K, \hat \phi) = 0
  \end{equation}
  So the second term of \eqref{eq:xxie3} vanishes. It remains to show
  \begin{eqnarray*}
    \pd{\mathcal U(F, \phi)}{\alpha}(\alpha, K, \hat \phi)
    &>& 0
  \end{eqnarray*}
  From \eqref{eq:xxie1.0}, it follows
  \begin{eqnarray*}
    \pd{\mathcal U(F, \phi)}{\alpha}(\alpha, K, \hat\phi)
    = \opd{\alpha}\E[u((1 - \hat\phi) e^r + \hat\phi e^X) \1{X < 0}]
  \end{eqnarray*}
  The function $u((1 - \hat\phi) e^r + \hat\phi e^X)$ is obviously increasing
  with $X$. It follows
  \begin{eqnarray*}
    && \pd{f(x; \alpha, K)}{\alpha} \\
    &=& \opd{\alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} \\
    &=&
    - {K^\alpha \over (K - x)^{\alpha + 1}}
    \left[
      \alpha
      \ln\left(
        1 - {x \over K}
      \right) - 1
    \right]
  \end{eqnarray*}
  It is easily checked that when $x < K(1 - e^{1/\alpha}) < 0$,
  \[
  \opd{\alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} < 0
  \]
  and when $K(1 - e^{1/\alpha}) < x < 0$,
  \[
  \opd{\alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} > 0
  \]
  This is the second case of lemma \ref{lemma:I}. So we have
  \begin{eqnarray*}
    \pd{\mathcal U(F, \phi)}{\alpha}(\alpha, K, \hat\phi) &>& 0
  \end{eqnarray*}
  As for $\pd{G(F)}{K}$, by the same argument as for
  $\pd{G(F)}{\alpha}$, it suffices to show
  \[
  \pd{\mathcal U(F, \phi)}{K}(\alpha, K, \hat\phi) < 0
  \]
  We have
  \[
  \pd{\mathcal U(F, \phi)}{K}(\alpha, K, \hat\phi)
  = \opd{K} \E \left[
    u((1 - \hat\phi) e^r + \hat\phi e^X) \1{X < 0}
  \right] 
  \]
  and
  \begin{eqnarray*}
    && \pd{f(x; \alpha, K)}{K} \\
    &=& \opd{K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} \\
    &=&
    -\alpha K^{\alpha - 1} {
      \alpha x + K
      \over
      (K - x)^{\alpha + 2}
    }
  \end{eqnarray*}
  Clearly, when $x < -\alpha/K$
  \[
  \opd{K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} > 0
  \]
  and when $-\alpha/K < x < 0$
  \[
  \opd{K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} < 0
  \]
  So by the 1st case of remark \ref{remark:I} we conclude
  \[
  \pd{\mathcal U(F, \phi)}{K}(\alpha, K, \hat\phi) < 0  
  \]
\end{proof}

@@

\subsubsection{log-utility function}
We first consider the case where the utility function is simply
$u(\cdot) = \ln(\cdot)$. This corresponds to the limiting case of
\eqref{eq:power_utility} as $\gamma \to 0$. Straightforward
computation gives
\begin{eqnarray*}
  && \mathcal U(F, \phi) \\
  &=&
  \int_{0}^{\infty}
  \left[
    \ln(C(x)) + (1 + b)\ln(C(-x))
  \right]
  f(x, \nu) dx
  + b \int_{0}^{q}
  \ln(C(x))
  f(x, \nu) dx \\
  &&
  - b \ln(C(q)) F(q, \nu)
\end{eqnarray*}
where $C(\cdot)$ is defined in \eqref{eq:xxie1}. Notice that
when $b = 0$, $\mathcal U(F, \phi)$  reduces to $\E u(C(X))$.
Also observe $\ln(C(x)C(-x))$ is an increasing function:
\begin{eqnarray*}
  \pd{C(x) C(-x)}{x}
  &=&
  \phi e^r (1 - \phi) (e^x - e^{-x}) > 0
\end{eqnarray*}
Thus by 1st case of remark \ref{remark:I} and lemma \ref{lemma:II},
$\pd{\mathcal U(F, \phi))}{\nu} < 0$ for all $\phi$. Hence all
investors guided by this utility function would be seeking the
smallest $\nu$, i.e. the heaviest tail in the market.

>>>>

\subsection{Pareto-distributed equity return}
In this section, we look into the situation when $X$ has a shifted
Pareto distribution. Assume
\begin{equation}
  \label{eq:X_distr}
  F(x) = \left\{
  \begin{array}{ll}
    p \left(
    {K \over K - x}
    \right)^\alpha & x \leq 0 \\
    1 - (1 - p) \left(
    {K' \over K' + x}
    \right)^\beta & x > 0
  \end{array}
  \right.
\end{equation}
Assume $\alpha, \beta > 1$, $0 < p < 1$ and $K > 0$.
We have
\begin{eqnarray}
  && \mathcal U F (\alpha, \beta, K, K', \phi) \nonumber \\
  &=&
  \alpha K^\alpha  p
  \int_{-\infty}^0
  u\left[ (1 - \phi) e^r + \phi e^x \right]
  {1 + b \over (K - x)^{\alpha + 1}} dx
  \nonumber \\
  &&
  \beta K'^\beta (1 - p)
  \int_{0}^\infty
  u\left[ (1 - \phi) e^r + \phi e^x \right]
  {1 + b \1{x < q} \over (K' + x)^{\beta + 1}} dx 
  - b u(\delta v) F(q)
  % &&
  % - b u(\delta v) \left[
  %   1 - (1 - p) {
  %     K'^\beta
  %     \over
  %     (K' + q)^\beta
  %   }
  % \right]
  \label{eq:xxie1.0}
\end{eqnarray}
\begin{theorem}
  \label{thrm:I}
  Suppose functional $G(\cdot)$ is given as \eqref{eq:G_functional},
  function $u(\cdot)$ is increasing and differentiable, then
  \[
  \pd{G(F)}{\alpha} > 0,
  \quad
  \pd{G(F)}{K} < 0
  \]
\end{theorem}
Since $u(v)$ given in \eqref{eq:xxie1.0} as a
functional of the distribution function of $X$ is increasing with the
lower tail index $\alpha$ and decreasing with $K$, there is a curve of
equal preference on the plane of $(\alpha, K)$. Moving along this
curve in the direction of increasing $\alpha$, one expects the values
of $K$ to increase too, i.e. the estimated values of $\alpha$ and $K$
should appear positively correlated. This is indeed the case in some
real stock return data, e.g. the ``Energy'' and ``Information Technology''
sectors of S\&P 500 companies. This is illustrated in figures
\ref{fig:Energy_alpha_K} and \ref{fig:Information_Technology_alpha_K}.
The confidence bands of the Hill estimates of $\alpha$ of these
sectors are shown in figures \ref{fig:Energy_lower},
\ref{fig:Consumer_Staples_lower} and
\ref{fig:Information_Technology_lower}.

\begin{proof}
  Let
  \begin{equation}
    \label{eq:xxie2}
    \hat \phi := \argmax_{0 < \phi \leq 1} \mathcal U(F, \phi)
  \end{equation}
  We have
  \[
  G(F) = \mathcal U(F, \hat\phi)
  \]
  It follows
  \begin{eqnarray}
    \pd{G(F)}{\alpha}
    &=&
    \pd{\mathcal U(F, \phi)}{\alpha}(\alpha, K, \hat\phi)
    + \pd{\mathcal U(F, \phi)}{\phi}(\alpha, K, \hat\phi)
    \pd{\hat\phi}{\alpha}(\alpha, K)
    \label{eq:xxie3}
  \end{eqnarray}
  The definition \eqref{eq:xxie2} implies $\forall \alpha > 1$
  \begin{equation}
    \label{eq:xxie4}
    \pd{\mathcal U(F, \phi)}{\phi}(\alpha, K, \hat \phi) = 0
  \end{equation}
  So the second term of \eqref{eq:xxie3} vanishes. It remains to show
  \begin{eqnarray*}
    \pd{\mathcal U(F, \phi)}{\alpha}(\alpha, K, \hat \phi)
    &>& 0
  \end{eqnarray*}
  From \eqref{eq:xxie1.0}, it follows
  \begin{eqnarray*}
    \pd{\mathcal U(F, \phi)}{\alpha}(\alpha, K, \hat\phi)
    = \opd{\alpha}\E[u((1 - \hat\phi) e^r + \hat\phi e^X) \1{X < 0}]
  \end{eqnarray*}
  The function $u((1 - \hat\phi) e^r + \hat\phi e^X)$ is obviously increasing
  with $X$. It follows
  \begin{eqnarray*}
    && \pd{f(x; \alpha, K)}{\alpha} \\
    &=& \opd{\alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} \\
    &=&
    - {K^\alpha \over (K - x)^{\alpha + 1}}
    \left[
      \alpha
      \ln\left(
        1 - {x \over K}
      \right) - 1
    \right]
  \end{eqnarray*}
  It is easily checked that when $x < K(1 - e^{1/\alpha}) < 0$,
  \[
  \opd{\alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} < 0
  \]
  and when $K(1 - e^{1/\alpha}) < x < 0$,
  \[
  \opd{\alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} > 0
  \]
  This is the second case of lemma \ref{lemma:I}. So we have
  \begin{eqnarray*}
    \pd{\mathcal U(F, \phi)}{\alpha}(\alpha, K, \hat\phi) &>& 0
  \end{eqnarray*}
  As for $\pd{G(F)}{K}$, by the same argument as for
  $\pd{G(F)}{\alpha}$, it suffices to show
  \[
  \pd{\mathcal U(F, \phi)}{K}(\alpha, K, \hat\phi) < 0
  \]
  We have
  \[
  \pd{\mathcal U(F, \phi)}{K}(\alpha, K, \hat\phi)
  = \opd{K} \E \left[
    u((1 - \hat\phi) e^r + \hat\phi e^X) \1{X < 0}
  \right] 
  \]
  and
  \begin{eqnarray*}
    && \pd{f(x; \alpha, K)}{K} \\
    &=& \opd{K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} \\
    &=&
    -\alpha K^{\alpha - 1} {
      \alpha x + K
      \over
      (K - x)^{\alpha + 2}
    }
  \end{eqnarray*}
  Clearly, when $x < -\alpha/K$
  \[
  \opd{K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} > 0
  \]
  and when $-\alpha/K < x < 0$
  \[
  \opd{K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} < 0
  \]
  So by the 1st case of remark \ref{remark:I} we conclude
  \[
  \pd{\mathcal U(F, \phi)}{K}(\alpha, K, \hat\phi) < 0  
  \]
\end{proof}
By numerically integration and optimization with respect to $\phi$,
the value of $\hat\phi$ can be obtained for given $u(\cdot)$ and given
values of $K, K', \alpha, \beta$. Figures \ref{fig:phi_hat_pareto5e-1}
and \ref{fig:phi_hat_pareto4} show the values of $\hat\phi$ for
$u(C) = -\frac{C^{-\xi}}{\xi}$ with $\xi = 1/2, 4$ respectively.
$\beta=\alpha$ and $K'=K$ in both figures.
\begin{figure}
  \begin{subfigure}[b]{0.5\linewidth}
    \includegraphics[width=\textwidth]{phi_hat_pareto5e-1.pdf}
    \caption{$\hat\phi$}
    \label{fig:phi_hat_pareto5e-1}
  \end{subfigure}
  \begin{subfigure}[b]{0.5\linewidth}
    \includegraphics[width=\textwidth]{phi_hat_pareto4.pdf}
    \caption{$\hat\phi$}
    \label{fig:phi_hat_pareto4}
  \end{subfigure}
  \caption{$\hat\phi$, the optimal equity allocation when the utility
    function is $u(C) = -{C^{-\xi} \over \xi}$ for $\xi = 1/2, 4$.
  }
\end{figure}

\begin{minipage}{0.5\linewidth}
  \includegraphics[width=\textwidth]{power_utilities.png}
\end{minipage}\hfill
\begin{minipage}{0.42\textwidth}
  As shown to the left, the power utility function with a small
  $\xi$ grows slower and saturates later than does one with a large
  $\xi$. It represents an agent that is more tolerant of low
  consumption and that seeks wealth more aggressively. In other
  words, he is less risk-avert than is one with a higher $\xi$.
  Such an agent will therefore invest more heavily on the equity as
  is shown in figures \ref{fig:phi_hat_pareto5e-1} and
  \ref{fig:phi_hat_pareto4}.
\end{minipage}
Certainly, it rarely happens in practice that the lower- and
upper-tail indices of an equity are equal. Thus, to verify our result
that $G(F)$ is increasing with $\alpha$ and decreasing with $K$, we
need to verify that the 1st integral of \eqref{eq:xxie1.0} with $\phi$
replaced by $\hat\phi$, i.e.
\[
I(\alpha, K) = 
  \alpha K^\alpha  p
  \int_{-\infty}^0
  u\left[ (1 - \hat\phi(\alpha, K)) e^r + \hat\phi(\alpha, K) e^x \right]
  {1 + b \over (K - x)^{\alpha + 1}} dx
\]
is increasing with $\alpha$ and decreasing with $K$. Figures
\ref{fig:preference_pareto5e-1} and \ref{fig:preference_pareto4}
show the values of the above integral when $\alpha$ and $K$ take
a range of values.
\begin{figure}[htb!]
  \begin{subfigure}[b]{0.5\linewidth}
    \includegraphics[width=\textwidth]{preference_pareto5e-1.pdf}
    \caption{$I(\alpha, K)$}
    \label{fig:preference_pareto5e-1}
  \end{subfigure}
  \begin{subfigure}[b]{0.5\linewidth}
    \includegraphics[width=\textwidth]{preference_pareto4.pdf}
    \caption{$I(\alpha, K)$}
    \label{fig:preference_pareto4}
  \end{subfigure}
  \caption{$I(\alpha, K)$ when the utility function is
    $u(C) = -{C^{-\xi} \over \xi}$ for $\xi = 1/2, 4$.
  }
\end{figure}

@@

We consider the case where the utility function is simply
$u(\cdot) = \ln(\cdot)$. This corresponds to the limiting case of
\eqref{eq:power_utility} as $\gamma \to 0$. Straightforward
computation gives
\begin{eqnarray*}
  && \mathcal U(F, \phi) \\
  &=&
  \int_{0}^{\infty}
  \left[
    \ln(C(x)) + (1 + b)\ln(C(-x))
  \right]
  f(x, \nu) dx
  + b \int_{0}^{q}
  \ln(C(x))
  f(x, \nu) dx \\
  &&
  - b \ln(C(q)) F(q, \nu)
\end{eqnarray*}
where $C(\cdot)$ is defined in \eqref{eq:xxie1}. Notice that
when $b = 0$, $\mathcal U(F, \phi)$  reduces to $\E u(C(X))$.
Also observe $\ln(C(x)C(-x))$ is an increasing function:
\begin{eqnarray*}
  \pd{C(x) C(-x)}{x}
  &=&
  \phi e^r (1 - \phi) (e^x - e^{-x}) > 0
\end{eqnarray*}
Thus by 1st case of remark \ref{remark:I} and lemma \ref{lemma:II},
$\pd{\mathcal U(F, \phi))}{\nu} < 0$ for all $\phi$. Hence all
investors guided by this utility function would be seeking the
smallest $\nu$, i.e. the heaviest tail in the market.
